Reconstructing Propositional Proofs in Type Theory

### 2017-10-24 paper/sections/introduction.tex
'.
[ paper ] last changes.
[ paper ] some messages added.
[ paper ] none inspiration .. :S

\section{Introduction}
\label{sec:introduction}

% Proof-reconstruction is a non-trivial task since it depends on the
% integration between two complex systems.
% We describe one proof-reconstruction approach between an automatic theorem
% prover (henceforth \ATP) and an interactive theorem prover (henceforth
% \ITP).

% The \ITP tools provide us the
% logic framework to check and validate the reply of the \ATPs since
% they allow us to define the formal language for the problems i.e.,
% operators, logic variables, axioms, and theorems.

@@

% Many approaches have been proposed and some tools have been
% implemented for proof-reconstruction in the last decades. These
% programs are relevant not only because it helps to spread their
% usage but they also increase the confidence of their users about
% their algorithms and their correctness (see, for example, bugs in
% \ATPs~\cite{Keller2013,Bohme2011,Fleury2014} and
% \cite{Kanso2012}).
% We mention some of these tools in Section~\ref{sec:related-work}.

% In this paper, we ported a subset of the \Metis' inference rules to
% \Agda, the propositional fragmented. The formalization of
% these rules allowed us to justify step-by-step \Metis' \TSTP derivations by
% translating automatically to \Agda by a proof-reconstruction tool built for
% this research named \Athena.
% This paper has been organized in the following way.
% In Section~\ref{ssec:CPL}, we introduce some
% notation and syntax of propositional logic.
% In Section~\ref{sec:metis-language-and-proofs}, we introduce the
% \Metis prover and its deduction system.
% In Section~\ref{sec:proof-reconstruction}, we present our
% approach to reconstruct proofs deliver by \Metis in type theory.
% Section~\ref{sec:conclusions} summarizes the main findings and includes a
% discussion of future work.
% Some related works to this paper are presented in Section~\ref{sec:related-work}.
% Finally, in appendix \textsc{A}, we provide a small example of the
% proof-reconstruction workflow.

>>>>

\section{Introduction (ESCRIBIENDO)}
\label{sec:introduction}

disciplina
en el tiempo
hilighting an import

The main challenge faced of reconstrucing proofs

the main disadavantage of X

One of the main obstacles to reconstruct proofs generated by
automatic provers consists of ...

However,

The performance of X is limited by ...

synopsys

Previous studies have reported proof-reconstruction for

Existing research recognizes the critical role by ...

What we known about X is largely based on ....

Few published studied have examined the role of

focus on a full description of the inference rules in the ATP logic.

No previous studies has given sufficient considerations to ...

This paper describes the design and implementation of X...

This study provides new insights into the automatic reasoning of \Metis prover.
Understanding the link between the ATP and ITP will help xxxx

The reader should bear in mind that the study is based on
reverse engineering.

DECIR QUE ES NOVEDOSO EN AGDA Y Q NO ESTA HECHO EN TEORIA DE TIPOS
SYNTATICA VS SEMANTICA.
APPLICACIONES

% We describe how to reconstruct proofs delivered by an automatic theorem prover
% (henceforth \ATP)
% and an interactive theorem prover (henceforth
% \ITP).

% One way to reconstruct proofs of \ATPs consists of using an \ITP.
% An \ITP system provide us the logic framework to check and validate
% the reply of the \ATPs since they allow us to define the formal language
 % for the problems \ie, operators, logic variables, axioms, and theorems.

% This task is non-trivial since it depends on the
% integration between two complex systems.

@@

 These
% programs are relevant not only because it helps to spread their
% usage but they also increase the confidence of their users about
% their algorithms and their correctness (see, for example, bugs in
% \ATPs~\cite{Keller2013,Bohme2011,Fleury2014} and
% \cite{Kanso2012}).
% We mention some of these tools in Section~\ref{sec:related-work}.

The recontruction of  proof can play an important role in addressing a issue
of ...
Moreover, verification is fast becoming a key instrument in automatic reasoning.

The type theory provides a useful account of how ...

No previous study has investigated to reconstruct proofs using a syntactical approach
in \Agda for any \ATPs.

In this paper, we ported a subset of the \Metis inference rules to
\Agda, the propositional fragmented. The formalization of
these rules allowed us to justify step-by-step \Metis derivations by
translating automatically to \Agda by a the \Athena tool.

This paper has been organized in the following way.
We mention in Section~\ref{sec:limitations-type-theory}
some limitations of the proof-reconstruction from the type theory
point of view. In Section~\ref{sec:metis-language-and-proofs}, we introduce the
\Metis prover and its deduction system.
In Section~\ref{sec:proof-reconstruction}, we show our
approach to reconstruct proofs deliver by \Metis.
Related work is described in Section~\ref{sec:related-work}.
Section~\ref{sec:conclusions} summarizes the main findings and suggest future work.

% Finally, in appendix \textsc{A}, we provide a small example of the
% proof-reconstruction workflow.

### 2017-10-24 paper/sections/metis-language-and-proofs.tex
[ paper ] fixing refs.
[ paper ] none inspiration .. :S

@@

\begin{table}[!ht]
\label{tab:agda-metis-table}
\caption{\Metis inference rules.}
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis rule} & \textbf{Purpose} &\textbf{Theorem}\\ \hline
    \texttt{canonicalize}
      &Normalization of the formula
      &\ref{thm:canonicalize}
    \\
    \texttt{clausify}
      &Performs clausification
      &\ref{thm:clausify}
    \\
    \hyperlink{atp-conjunct}{\texttt{conjunct}}
      &Takes a formula from a conjunction
      &\ref{thm:conjunct}
    \\
    % \texttt{negate}
    %   &Negate the formula
    %   &\ref{thm:negate}
    % \\
    \texttt{resolve}
      &A general form of the resolution theorem
      &\ref{thm:resolve}
    \\
    \texttt{simplify}
      &Simplify definitions and theorems
      &\ref{thm:simplify}
    \\
    \texttt{strip}
      &Strip a goal into subgoals
      &\ref{thm:strip}
    \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\end{table}

>>>>

We note a lack of clarity in defining the differences between the rules used in \TSTP derivations and the rules described on the documentation in
the Metis description~\cite{hurd2003first}.
These rules are not the same as we expected from the Fig.~\ref{fig:metis-inferences}.

AGREGAR el sitio de metis official.!!

@@

\begin{table}[!ht]

\caption{\Metis inference rules.}
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \label{tab:agda-metis-table}
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis rule} & \textbf{Purpose} &\textbf{Theorem}\\ \hline
    \texttt{canonicalize}
      &Normalization of the formula
      &\ref{thm:canonicalize}
    \\
    \texttt{clausify}
      &Performs clausification
      &\ref{thm:clausify}
    \\
    \hyperlink{atp-conjunct}{\texttt{conjunct}}
      &Takes a formula from a conjunction
      &\ref{thm:conjunct}
    \\
    % \texttt{negate}
    %   &Negate the formula
    %   &\ref{thm:negate}
    % \\
    \texttt{resolve}
      &A general form of the resolution theorem
      &\ref{thm:resolve}
    \\
    \texttt{simplify}
      &Simplify definitions and theorems
      &\ref{thm:simplify}
    \\
    \texttt{strip}
      &Strip a goal into subgoals
      &\ref{thm:strip}
    \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\end{table}

### 2017-10-24 paper/sections/proof-reconstruction/strip.tex
[ paper ] none inspiration .. :S

\Metis proves each subgoal in the same order above from left to right
in (\ref{eq:strip-example}).
Besides the fact \Metis does not explicit in \TSTP derivation
how it uses the proofs of subgoals, we know using the elimination
of the implication, we get the proof of the conjecture.
\end{myexample}

>>>>

\Metis proves each subgoal in the same order above from left to right
in (\ref{eq:strip-example}).
So far, very little attention has been paid to the role of \strip rule.
in \TSTP derivation since \Metis does not make explicit the way
it uses the subgoals to prove the conjecture.
\end{myexample}
